---
title: "Stochastic Control & Optimization Project 2"
author: "Anthony Garino, Davis Townsend, Dallas Griffin, Ryan Maas"
date: "February 16, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Stochastic Control & Optimization Project 2


Below is the code prof provided to do initial data cleaning and created price matrix, shares outstanding matrix, and vector of the tickers
```{r}
# You need to set your own directory

# read in the data
data = read.csv("N100StkPrices.csv", header = TRUE)


# clean up data
data = na.omit(data)
ticker = data$TICKER

# spun off MDLZ
delete = seq(1, dim(data)[1])[ticker == "MDLZ"]
data = data[-delete, ]

date = apply(as.matrix(data$date), MARGIN = 1, FUN = "toString")
date = as.Date(date, "%Y%m%d")
ticker = data$TICKER
price = data$PRC
shares = data$SHROUT


# Accounting for changes in ticker names


# KFT changed to KRFT in Oct 2012.
ticker[ticker == "KFT"] = "KRFT"

# SXCI changed to CTRX in Jul 2012.
ticker[ticker == "SXCI"] = "CTRX"

# HANS changed to MNST in Jan 2012.
ticker[ticker == "HANS"] = "MNST"



# convert prices to a matrix, arranged by rows of dates and columns of tickers
unique_dates = sort(unique((date)))
unique_tickers = sort(unique(ticker))

priceMat = matrix(NA, length(unique_dates), length(unique_tickers))
sharesMat = matrix(0, length(unique_dates), length(unique_tickers))



for (i in 1:length(unique_tickers)) {
  tic = unique_tickers[i]
  idx = is.element(unique_dates, date[ticker == tic])
  
  priceMat[idx, i] = price[ticker == tic]
  sharesMat[idx, i] = shares[ticker == tic]
}

rownames(priceMat) = as.character(unique_dates)
rownames(sharesMat) = as.character(unique_dates)

rm(list = c("data", "delete", "i", "idx", "price", "shares", "tic", "ticker", "date"))

unique_tickers

```


1. Below is the code to generate the daily returns matrix for each stock:
(important to note this is put into percentage terms so 1.0 means 1% daily stock return)
```{r}

#create empty matrix to populate with the results
returnsMatrix = matrix(nrow=250, ncol=100)

#loop through every row and every column, if not in the first row calculate 
#daily return by the formula:
# ((closing price - prev close price)/prev close price)*100
for(i in 1:nrow(priceMat)) {
  for(j in 1:ncol(priceMat)) {
    if (i>1){
      returnsMatrix[i,j] = ((priceMat[i,j] - priceMat[i-1,j])/priceMat[i-1,j])*100
    }
    
  }
}

head(returnsMatrix)

```

Calculate correlation matrix for daily returns
```{r}

constructFund = function(rho, q, priceMat, sharesMat, unique_tickers, unique_dates){
  
  library(lpSolveAPI)
  
  n = dim(priceMat)[2]
  
  # if there's n = 100 stocks, theres 100*100+100 = 10100 decision variables
  # 10000 x~ij~ values
  # 100 y~j~ values
  # The number of constraints depends on n; there are n^2 + n + 1 constraints
  
  LP=make.lp(0,n^2+n)
  set.objfn(LP, c(as.vector(stockCorr), rep(0,n)))
  lp.control(LP,sense='max')
  set.type(LP, (1:n^2+n), "binary") #forces all variables to be binary
  
  # constraint x~ij~ <= y~j
  for (i in 1:n^2){
    if (i%%n != 0){
      add.constraint(LP,c(rep(0,i-1),1,rep(0,n^2-i),rep(0,i%%n-1),-1,rep(0,n-i%%n)), "<=", 0)
    }else{
      add.constraint(LP,c(rep(0,i-1),1,rep(0,n^2-i),rep(0,n-1),-1), "<=", 0)
    }
  }
  
  # constraint sum(x~ij~) = 1
  for (i in 1:n){
    add.constraint(LP,c(rep(0,(i-1)*n),rep(1,n),rep(0,(n-i)*n),rep(0,100)), "=", 1)
  }
  
  # contraint sum(y~j~) = q
  add.constraint(LP,c(rep(0,n^2), rep(1,n)), "=", q)
  
  # solve linear program
  solve(LP)
  #get.objective(LP)
  solution = get.variables(LP)
  x = solution[1:10000]
  y = tail(solution,100)
  
  x_mat = matrix(x,nrow = 100,ncol = 100, byrow=TRUE)
  
  # gets the share price and number of shares from the last date - 12/31
  marketValue = tail(sharesMat,1)*tail(priceMat,1)
  totalMarket = sum(marketValue)
  
  w = array()
  
  for (i in 1:n){
    w = c(w, sum(x_mat[,i]*marketValue)/totalMarket)
  }
  w = tail(w,100)
  
  return(w)
}

```


```{r}

stockCorr = cor(returnsMatrix, y=NULL, use="complete.obs")

q = 25

weights = constructFund(stockCorr, q, priceMat, sharesMat, unique_tickers, unique_dates)

```